---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# matchsurv

The goal of matchsurv is to estimate the cumulative excess risk for exposed individuals when matched survival data are available. Excess risk regression models are available on both hazard and cumulative incidence scale. More info in `vignette("matchsurv")`.


## Installation

You can install matchsurv from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("cribosch/matchsurv")
```

and call the package with:

```{r callpackage, results='hide', warning=FALSE}
library(matchsurv)

## other useful packages
library(timereg)
library(geepack)
library(ggplot2)
```

The package works with matched cohort data where the outcome on study is time-to-event. 
For each exposed individual we have a defined number of unexposed individuals, matched according to some relevant factors (the number of unexposed individuals per exposed can be different). To simulate data:

```{r datasim hazard}
dhaz<-sim.data.MatchH(5000,5)
head(dhaz,10)
```

Or when working with cumulative incidence functions:

```{r datasim cif}
dcif<-sim.data.MatchCR(1000,5)
head(dcif, 10)
```

Further info about the function options are available through `vignette()` or `example()`.

This is a basic example which shows you how to:

1) estimate the excess risk 
2) visualize your results - coefficient estimates and cumulative excess 

in both the two settings.

## Model estimate

### Excess hazard 

Data set up

```{r compdata, results='hide'}
example("compdata")
```

Model estimate

```{r modelest, results='hide'}
example("matchpropexc")
```

### Excess cumulative incidence

Data set up

```{r}
example("compcomp")

```

Model estimate based on GEE function `geepack::geese()`

```{r}
tp<-c(0.5,1,2,5,10,15,25)
setdcif1<-compcomp(Event(time=FALSE,time2=time,cause=cause)~X1+X2,
data=dcif, cluster=i, idControl=j, time.points=tp, cens.formula=NULL, event=1)

exc.cif.mod1<-geese(Rt~-1+factor(h)+X1+X2,
                    data=setdcif1,
                    family="gaussian", #error distribution
                    mean.link = "log", #link function for Rt
                    corstr="independence", #correlation structure
                    id=clust.num, #cluster vector
                    weights=weights #censoring weights
)
```


## Results

### Excess hazard

To visualize the coefficient estimates: `summary(model)`

To estimate the cumulative baseline excess hazard:

```{r cumhaz, results='hide'}
example("exccumhaz")
```

To plot the cumulative baseline excess hazard:

*Note: if your model has strata, you can chose which strata to plot (option: `stratas=`, followed by the number of the strata, the first one is number 0). You can also decide to show the relative survival (option: `relsurv=TRUE`).*

```{r plots, results='hide', fig.show='hide', eval=FALSE}
example("excplot")
```

### Excess cumulative incidence

To visualize the estimated effects: 

```{r}
example("ecif.coef")
```

To predict the excess cumulative incidence for different covariate values:

```{r}
example("ecif.pred")
```

Further info on plotting the results available in `vignette(matchsurv)`

