---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# matchsurv

The goal of matchsurv is to estimate the cumulative excess risk for exposed individuals when matched survival data are available. 


## Installation

You can install matchsurv from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("cribosch/matchsurv")
```

## Example

### matched survival data

```{r callpackage, results='hide', warning=FALSE}
library(matchsurv)
```
For each exposed individual we have a defined number of unexposed individuals, matched according to some relevant factors (the number of unexposed individuals per exposed can be different). 

Here it is an example of the data:
```{r datasim}
d<-data.sim(5000,5)
head(d,10)
```

`matchsurv::data.sim` let you simulate some matched survival data; `competing=TRUE` will let you chose for a competing risk setting; when `nullmod=TRUE` no covariates are simulated.

This is a basic example which shows you how to:

1) estimate the proportional excess model 
2) visualize your results - coefficient estimates and cumulative excess hazard plot

### model estimate

First you need to set up your data in order to estimate the model

```{r compdata}
names(d)
setd<-compdata(Surv(time, status)~x+z+cc, data=d, idControl = j, cluster=id)
names(setd)
```

`strata()` are not needed in here; you'll choose after how to specify your model. It is a good idea to use all the possible covariates at this step. New variables will be created, you'll need them in the following step.

Then you can estimate your model:

```{r modelest}
m <- matchpropexc(Surv(exit,status)~strata(z)+factor(x),cluster=cluster, idControl=unexp.subj, weight=weight,data=setd)
```

## model results

To visualize the coefficient estimates

```{r summary}

summary(m)

```

To estimate the cumulative baseline excess hazard:

```{r cumhaz}

cumhaz <- exccumhaz(m) #it's a list because of strata
cumhaz <- exccumhaz(m, time=seq(0,30,5)) #you can chose at which time-points to show the estimates

```

To plot the cumulative baseline excess hazard:

*Note: if your model has strata, you can chose which strata to plot (option: `stratas=`, followed by the number of the strata, the first one is number 0). You can also decide to show the relative survival (option: `relsurv=TRUE`).*

```{r plots, fig.height=10, fig.width=8}
par(mfrow=c(3,2))
excplot(m, se=TRUE, col=c("green","blue"), main="with polygon CI") 
excplot(m, se=TRUE, time=seq(0,30,1), main="at specific time-points") 
excplot(m, se=TRUE, relsurv=TRUE, main="relative surv.") 
excplot(m, se=TRUE, polygon=FALSE, main="with CI - no polygon") 
excplot(m, se=FALSE, main="No CI") 
excplot(m, se=TRUE, stratas=1, main="plot the second strata") 
```

