---
title: "matchsurv: excess risk estimation in matched cohort studies"
author: "Cristina Boschini"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{matchsurv: excess risk estimation in matched cohort studies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

requireNamespace("devtools")
```

`matchsurv` provides  `R` functions to estimate excess risk regression models on both the hazard and the cumulative incidence scale. 

# Installation
You can install `matchsurv` from `github` with:

```{r, echo=TRUE, results="hide",eval=FALSE}
# install.packages("devtools")
devtools::install_github("cribosch/matchsurv")
```

### libraries

```{r}
library(matchsurv)

#other useful packages 
library(timereg)
library(geepack)
library(ggplot2) # for the plot
```

These functions work with matched cohort studies where the outcome is *time-to-event*. The excess risk can be estimated in a survival setting on the hazard scale and in a cometing risk setting also in terms of cumulative incidence. Data can be simulated in both the two settings.

# Excess hazard model

## Data simulation

We simulate matched cohort data in the survival setting with 5000 exposed subjects and 5 matched unexposed subjects for each exposed. 

```{r}
haz.data<-sim.data.MatchH(5000,5)
head(haz.data)
```

The option`competing=TRUE` of `sim.data.MatchH` functions allows to estimate data in a competing risk setting where the excess risk is defined in terms of cause specific hazard. When `nullmod=TRUE`, the excess risk does not depend on covariates, so no covariates are simulated.  

## Model estimation

To get the model estimates we first need to set up the data correctly; both the two approaches work on the exposed-unexposed pair information. 

```{r}
set.hazd<-compdata(Surv(time, status)~x+z+cc, clust=id, idControl=j, data=haz.data)
head(set.hazd)
```

We suggest to set all the possible coviariates in the perdictor part of the formula; this is not the estimated model yet. The dataset we obtain with `compdata` will be used in the model estimation. 

To estimate the model, run:

```{r}

exc.haz.mod<-matchpropexc(Surv(entry,exit,status)~strata(z)+factor(x), data=set.hazd)

```

Note that it is possible to specify possible strata or interactions. 

## Results

To visualise the estimated effects:

```{r}

summary(exc.haz.mod)

```

To estimate the cumulative baseline excess hazard:

```{r}
exccumhaz(exc.haz.mod, time = seq(0,25,5))
```

It is possible to specify a vector of timepoints where to estimate the cumulative baseline excess hazard. 
To plot the estimated cumulative hazard, we can use the function `excplot`; the option `stratas=` allows to visualize only some strata (strata are identified with a number, 0 refers to the first stratum). The option `relsurv=TRUE` plots the relative survival estimated through the excess risk model. 

```{r}
excplot(exc.haz.mod)
```

A better overview of the function `excplot()` is given by `example("excplot")`.

# Excess cumulative incidence model 

## Data simulation

Matched cohort data in a competing risk setting with two possible causes of failure are generated:

```{r}

cif.data<-sim.data.MatchCR(1000,5)
head(cif.data)
```


## Model estimate

### data set up

```{r}
tp<-c(0.5,1,2,5,10,15,25) # define vector of time points
set.cifd<-compcomp(Event(time=FALSE,time2=time,cause=cause)~X1+X2,
                   data=cif.data,
                   cluster=i, #cluster ID
                   idControl=j, #subject ID
                   time.points = tp, #time points
                   cens.formula = NULL, #censoring weight model
                   event=1 #event of interest
)
head(set.cifd)
```


### GEE 

```{r}
exc.cif.mod<-geese(Rt~-1+factor(h)+X1+X2,
                data=set.cifd,
                family="gaussian", #error distribution
                mean.link = "log", #link function for Rt
                corstr="independence", #correlation structure
                id=clust.num, #cluster vector
                weights=weights #censoring weights
                )

```


## Results 

### predicted excess CIFs

```{r}

af<-paste0("-1+factor(h)+X1+X2") #model formula
newd<-data.frame(expand.grid(h=tp,X1=c(0,1),X2=c(0.8,1.5,2.5))) # newdata
# define the different subjects for whom the excess risk is predicted
strata.levels<-factor(1:6, levels=1:6, 
                      labels =paste0(rep("X1=",6),
                                     expand.grid(X1=c(0,1),X2=c(0.8,1.5,2.5))[,1],
                                     rep(", X2=",6),
                                     expand.grid(X1=c(0,1),X2=c(0.8,1.5,2.5))[,2]))

pred.exc.cif<-ecif.pred(exc.cif.mod,times = tp,dataset = newd, 
                     formula = af, strata.levels = strata.levels)
head(pred.exc.cif,8)

```


### the plot


```{r}
p<-ggplot()
plot.exc.cif<-p+
  geom_step(aes(x=time, y=lower.ci, alpha=0.1), data=pred.exc.cif,lty="dotted", size=0.7)+
  geom_step(aes(x=time, y=upper.ci, alpha=0.1), data=pred.exc.cif,lty="dotted", size=0.7)+
  geom_step(aes(x=time, y=cif),col="black", data=pred.exc.cif, size=0.8)+
  scale_x_continuous(name="Time since entry",limits = c(0,30),breaks = seq(0,30,5), labels=seq(5,35,5))+
  scale_y_continuous(name="Excess risk")+ 
  facet_wrap(~strata,ncol = 2)+
  geom_abline(slope = 0,intercept = 0, size=0.2)+
  theme(legend.position = 'none');plot.exc.cif
```

### estimate effects

```{r}

ecif.coef(exc.cif.mod,times = tp, link = "log")

```



<!-- ## model estimate (cause 2) ----- -->
<!-- set.data2<-compcomp(Event(time=FALSE,time2=time,cause=cause)~X1+X2, -->
<!--                 data=data, -->
<!--                 cluster=i, #cluster ID -->
<!--                 idControl=j, #subject ID -->
<!--                 time.points = tp, #time points -->
<!--                 cens.formula = NULL, #censoring weight model -->
<!--                 event=2 #event of interest -->
<!--                 ) -->

<!-- exc.mod2<-geese(Rt~-1+factor(h), #no covariate effects -->
<!--                 data=set.data2, -->
<!--                 family="gaussian", #error distribution -->
<!--                 mean.link = "log", #link function for Rt -->
<!--                 corstr="independence", #correlation structure -->
<!--                 id=clust.num, #cluster vector -->
<!--                 weights=weights #censoring weights -->
<!--                 ) -->

<!-- af<-paste0("-1+factor(h)") -->
<!-- newd<-data.frame(h=tp) #no covariates -->
<!-- pred.exc2<-ecif.pred(exc.mod2,times = tp,dataset = newd, -->
<!--                      formula = af) -->
<!-- head(pred.exc2,8) -->

<!-- p<-ggplot() -->
<!-- plot.exc2<-p+ -->
<!--   geom_step(aes(x=time, y=lower.ci, alpha=0.1), data=pred.exc2,lty="dotted", size=0.7)+ -->
<!--   geom_step(aes(x=time, y=upper.ci, alpha=0.1), data=pred.exc2,lty="dotted", size=0.7)+ -->
<!--   geom_step(aes(x=time, y=cif),col="black", data=pred.exc2, size=0.8)+ -->
<!--   scale_x_continuous(name="Time since entry",limits = c(0,30),breaks = seq(0,30,5), labels=seq(5,35,5))+ -->
<!--   scale_y_continuous(name="Excess risk")+  -->
<!--   geom_abline(slope = 0,intercept = 0, size=0.2)+ -->
<!--   theme(legend.position = 'none');plot.exc2 -->

